Online Verification of Commutativity

### 2020-09-08 acmPaper/paper.tex
Finished first pass

There are several examples of reasonably complicated transformation graphs that we can pick from.  Gator includes several graphics examples as part of its examples package; for this evaluation, we looked at the \textit{phong}, \textit{reflection}, and \textit{shadow map} examples, sample images of which can be seen in Appendix~\ref{sec:gator-appendix}\xxx[as]{Either add text to appendix B or just reference a figure}.

We implemented a system for interfacing between the optimal set path checker (Algorithm 4) in the open-source implementation of Gator.
Using this algorithm, we found a bug in the reflection model, where a matrix was not being divided correctly by its determinant \xxx[dg]{List other things I find, if anything}\xxx[ak]{I think it may be nice to add more detail for at least one bug (code, what was the program doing, what the graph/conflict was, time taken, etc.)}.

@@

\section{Evaluation}
\label{sec:evaluation}
We compare performance of the following path checking algorithms:
(1) the na\"{i}ve baseline,
(2) the less na\"{i}ve two-flip baseline,
(3) the batch baseline,
(4) \ref{algo_online_polynomial}, and
(5) \ref{algo_online_minimal}.
\xxx[as]{Again, the reader probably needs a reminder about what the latter ones are.}
The two aspects we look at are time for response and size of response set (smaller sets---tighter output results---would mean less calls to the oracle).
We use randomly generated graphs for benchmarking, and vary their size.
Given a graph and a new edge, we time how long it takes for an algorithm to return  the set of pairs that need to be verified.
All computation was performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.

The na\"{i}ve baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
The batch algorithm brings timing down to seconds.
However time cost still does not scale very well, with computation for a graph with 14 nodes and 0.4 density taking hours.
Our implementation does not memoize the construction of the vector and matrix representation of paths in GF(2).
Profiling reveals that this vector construction step is called many times and is chiefly responsible for the high time cost.
\ref{algo_online_polynomial} performs only slightly better.
Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude, and runs in milliseconds for small graphs.
All implementations are sensitive to density, performing better when density is very low or very high.

@@

@@

Algorithms scale as expected with time, as seen in Figure~\ref{fig:timeVsSize}.
\xxx[as]{The plots are weirdly tiny on my machine... any chance your plotting tool can output PDF images?}
Both \ref{algo_online_minimal} and \ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance.
\xxx[as]{\ref{algo_minimal_spanning_set} does not appear in the figure.}
The na\"{i}ve baseline as well as the two flip tolerant baseline display quick growth.
The batch algorithm also grows fast, though not as much as the online checking baselines.

We define density to be the ratio of the number of edges in the graph to the total possible number of edges ($|V|^2$, where $|V|$ is the number of nodes).
Run time relates to the density of edges in the input graph.
The degree of the effect differs with the algorithms, as Figure~\ref{fig:timeVsSize} shows.
Generally, denser graphs entail longer computation time.
For the batch algorithm we use lower densities since the input graph must be acyclic. This puts an upper bound on density that approaches 0.5 in large graphs.
% Our proposed algorithms, ~\ref{algo_minimal_spanning_set} and ~\ref{algo_online_polynomial} are less affected because they aggressively remove redundancies.

>>>>

There are several examples of reasonably complicated transformation graphs that we can pick from.  Gator includes graphics examples as part of its examples package; for this evaluation, we looked at the \textit{phong}, \textit{reflection}, and \textit{shadow map} examples, sample images of which can be seen in Appendix~\ref{sec:gator-appendix}\xxx[as]{Either add text to appendix B or just reference a figure}.

We implemented a system for interfacing between the optimal set path checker (Algorithm 4) in the open-source implementation of Gator.
The system was tested with intentional bugs, of which it found them all, although no "real" bugs were found.  The graphs used were of size 5 or less; for graphs of this size, the checker was able to run in real time with no noticeable loss of frames.  Since the program is running at 60 frames per second, the checker was running at a rate faster than .01 seconds.

@@

\section{Evaluation}
\label{sec:evaluation}
We compare performance of the following path checking algorithms:
(1) the na\"{i}ve baseline,
(2) the less na\"{i}ve two-flip baseline,
(3) the batch baseline,
(4) \ref{algo_online_polynomial}, the non-minimal polynomial-time algorithm, and
(5) \ref{algo_online_minimal}, the minimal set finding algorithm.
\xxx[as]{Again, the reader probably needs a reminder about what the latter ones are.}
The two metrics we evaluate are time for response and size of response set (smaller sets---tighter output results---would mean less calls to the oracle).
We use randomly generated graphs of varying size: given a graph and a new edge, we time how long it takes for an algorithm to return the set of pairs that need to be verified.
All computations were performed on a Macbook Pro 2015, 2.9 GHz Dual-Core Intel Core i5.

@@

The na\"{i}ve baseline performs poorly, taking well over a thousand seconds for even small graphs of 10 nodes.
While the batch algorithm improves on this, it still does not scale very well, with computation for a graph with 14 nodes and 0.4 density taking hours.
Our implementation does not memoize the construction of the vector and matrix representation of paths in GF(2); profiling indicates that this construction is a major factor in the high time cost for these algorithms.
\ref{algo_online_polynomial} performs only slightly better than the batch algorithm.
Surprisingly, the optimal set algorithm cuts time cost by several orders of magnitude, and runs in milliseconds for small graphs.
All implementations are sensitive to density, performing better when density is low.

@@

Figure~\ref{fig:timeVsSize} shows that the algorithms time scales with size, as expected.
\xxx[as]{The plots are weirdly tiny on my machine... any chance your plotting tool can output PDF images?}
Both \ref{algo_online_minimal} and \ref{algo_minimal_spanning_set} exhibit graphs that are polynomial in appearance.
\xxx[as]{\ref{algo_minimal_spanning_set} does not appear in the figure.}
The na\"{i}ve baseline as well as the two flip tolerant baseline display quick growth.
The batch algorithm also grows fast, though not as much as the online checking baselines.

We define density to be the ratio of the number of edges in the graph to the total possible number of edges (which is $|V|^2$, where $|V|$ is the number of nodes).
Run time relates to the density of edges in the input graph.
The degree of the effect differs with the algorithms, as Figure~\ref{fig:timeVsSize} shows.
Generally, denser graphs entail longer computation time.
For the batch algorithm we use lower densities since the input graph must be acyclic. This puts an upper bound on density that approaches 0.5 in large graphs.
% Our proposed algorithms, ~\ref{algo_minimal_spanning_set} and ~\ref{algo_online_polynomial} are less affected because they aggressively remove redundancies.
